MS Prelim Data K23

### 2022-03-03 K23_prelim_data_started_20220225.Rmd
updated with phq9 stuff, didn't work

    *MS disease and depression network overlap:
      - Ms disease within depression network (35%) versus outside (17%), p < 2.2* 10^-16
      - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
      - MS disease within depression 5% network (30%) versus outside (17%), p < 2.2 * 10^-16
      - Linear Regressions (essentially matching t tests above): 
         +depressed v healthy within dep network in (p = 0.061); 
         +dep vs healthy within dep 10% (p = 0.049);
         +dep vs healthy within non dep network (p = 0.037)
      - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis. Suggests a final possible common pathway? Shared biology? This may also in part explain why rates of depression in MS are much higher than rates of depression in other autoimmune conditions. 
    *Depression by individual fascicle:
      - Nothing corrects at this time, unfortunately. 
      - P<0.05, uncorrected findings (* indicates association network):
          -AF_R	0.01168289	*		
          -FAT_R	0.01828607 *			
          -SLF2_R	0.01901378 *		
          -VOF_R	0.04741569 *			
          -CPT_F_R	0.02887163 			
          -CPT_O_L	0.02238344			
          -CPT_P_R	0.02053104			
          -CS_S_R	0.03783654			
          -DRTT_R	0.00695066			
          -ML_R	0.02511932	
    Fascicle loss x PHQ9:
     - future analysis?

@@

All the mixed models are below
```{r age_effects_indiv_fascicles_mm}

>>>>

    *MS disease and depression network overlap:
      - Ms disease within depression network (35%) versus outside (17%), p < 2.2* 10^-16
      - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
      - MS disease within depression 5% network (30%) versus outside (17%), p < 2.2 * 10^-16
      - Linear Regressions (essentially matching t tests above): 
         +depressed v healthy within dep network in (p = 0.061); 
         +dep vs healthy within dep 10% (p = 0.049);
         +dep vs healthy within non dep network (p = 0.037)
      - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis. Suggests a final possible common pathway? Shared biology? This may also in part explain why rates of depression in MS are much higher than rates of depression in other autoimmune conditions. 
    *Depression by individual fascicle:
      - Nothing corrects at this time, unfortunately. 
      - P<0.05, uncorrected findings (* indicates association network):
          -AF_R	0.01168289	*		
          -FAT_R	0.01828607 *			
          -SLF2_R	0.01901378 *		
          -VOF_R	0.04741569 *			
          -CPT_F_R	0.02887163 			
          -CPT_O_L	0.02238344			
          -CPT_P_R	0.02053104			
          -CS_S_R	0.03783654			
          -DRTT_R	0.00695066			
          -ML_R	0.02511932	
    Fascicle loss x PHQ9:
     - Nothing significant by individual fascicle, within association network, by fascicle within the depression network
     - similarly, no significant findings looks at means of all fascicle loss, within dep network, 10% and 5% and non-depression network

@@

```{r phq9_correlation_fascicles}

####### NOTHING SIGNIFICANT!!!! ########3
#isolate out only the depressed group and healthy group
with_phq9 <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0) %>%
  filter(!is.na(PHQ.9)) %>%
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#lm
fascicle_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
})
names(fascicle_lm) <- fascicle_names

#anova
fascicle_anova <- lapply(fascicle_lm, anova)

#fdr corrected
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
print(fascicle_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#uncorrected
fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
fascicle_anova_p05_unc <- fascicle_anova_p[!is.na(fascicle_anova_p) & fascicle_anova_p < 0.05]
print(fascicle_anova_p05_unc)

######### overall fascicle proportion affected - nothing is significant
overall_prop_affected <- lm(total_fascicle_prop_lost~PHQ.9, data=with_phq9)
overall_prop_affected_depnet <- lm(total_fascicle_prop_lost_indepnet~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_10percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_5percent <- lm(total_fascicle_prop_lost_indepnet_5_percent~PHQ.9, data=with_phq9)
overall_prop_affected_nondep <- lm(total_fascicle_prop_lost_nondep_network~PHQ.9, data=with_phq9)
```

```{r phq9_regressions_just_dep_group}
##### NOTHING SIGNIFICANT ########
##################################
## Just within depression group ##
##################################
#isolate out only the depressed group and healthy group
with_phq9 <- df_demo_and_fascicles %>%
  filter(depGroupVar == 2) %>%
  filter(!is.na(PHQ.9)) %>%
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()
#lm
fascicle_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
})
names(fascicle_lm) <- fascicle_names

#anova
fascicle_anova <- lapply(fascicle_lm, anova)

#fdr corrected
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
print(fascicle_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#uncorrected
fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
fascicle_anova_p05_unc <- fascicle_anova_p[!is.na(fascicle_anova_p) & fascicle_anova_p < 0.05]
print(fascicle_anova_p05_unc)

######### overall fascicle proportion affected - nothing is significant
overall_prop_affected <- lm(total_fascicle_prop_lost~PHQ.9, data=with_phq9)
overall_prop_affected_depnet <- lm(total_fascicle_prop_lost_indepnet~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_10percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_5percent <- lm(total_fascicle_prop_lost_indepnet_5_percent~PHQ.9, data=with_phq9)
overall_prop_affected_nondep <- lm(total_fascicle_prop_lost_nondep_network~PHQ.9, data=with_phq9)

#visreg
#sapply(fascicle_lm, visreg)
```
All the mixed models are below
```{r age_effects_indiv_fascicles_mm}
